---
title: "Migrating from blogdown/Hugo to jekylldown + al-folio"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Migrating from blogdown/Hugo to jekylldown + al-folio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

[al-folio](https://github.com/alshedivat/al-folio) is the de-facto
reference theme for academic sites -- publications generated from BibTeX,
news, CV, and teaching pages as structured data. Hugo has no equivalent,
which is the main reason to move a blogdown site to Jekyll. This vignette
walks through that migration with jekylldown, keeping your posts in
R Markdown throughout.

al-folio is the vignette's focus, but `migrate_hugo()` targets any of
the supported themes -- `theme = "chirpy"`, `"minimal-mistakes"` or
`"minima"` work the same way, with pages, navigation, avatar and social
profiles placed in each theme's own convention (Chirpy `_tabs/`,
Minimal Mistakes `_pages/` + `_data/navigation.yml` + the `author:`
sidebar, ...). Steps 1, 2 and 4 apply to all of them; step 3 is
al-folio-specific.

## Before you start

1. **Work on a copy.** `migrate_hugo()` never writes into the source
   site, but be conservative anyway: leave your production repository
   untouched until the new site builds and looks right.
2. **Diagnose your toolchain:**

```{r}
jekylldown::check()
```

A local Ruby/Jekyll is *optional* for this migration: knitting posts only
needs R, and al-folio ships GitHub Actions workflows that build the site
remotely. You only need Jekyll (plus `bundle_install()`, below) for local
previews.

## Step 1 -- convert the site

```{r}
library(jekylldown)

migrate_hugo(
  from  = "~/path/to/my-hugo-site",   # blogdown/Hugo site (read-only)
  to    = "~/path/to/my-new-site",    # created; must not exist yet
  theme = "al-folio",                 # fetched from GitHub (network needed)
  only_referenced = TRUE,             # skip static files nothing links to
  publications = "bib",               # the default; papers.bib from the DOIs (see below)
  theme_color = "red"                 # al-folio accent color (see below)
)
```

The Hugo config (`config.toml`/`.yaml`) is read first and drives the
migration. What gets converted automatically:

* **Posts** under `content/post/` (also `posts/`, `blog/`) are renamed to
  Jekyll's `YYYY-MM-DD-slug` convention. `.Rmd`/`.Rmarkdown` sources go
  to `_source/` and will be re-knitted by jekylldown; plain `.md` posts
  go straight to `_posts/`. Filename slugs are kept, so post URLs are
  preserved.
* **Front matter** (YAML or TOML) becomes Jekyll YAML: `layout: post` is
  added, `draft: true` becomes `published: false`, `aliases` becomes
  `redirect_from` (enable the `jekyll-redirect-from` plugin if you use
  it), and Hugo-only keys are dropped.
* **Page bundle resources** (images next to `index.Rmd`) are copied to
  `assets/img/posts/<post>/` and relative image links are rewritten.
* **Menu pages.** Every page referenced by the Hugo menu
  (`[[menu.main]]`) -- `content/publications.md`, `talks.md`, ... -- becomes
  an al-folio page in `_pages/` with the same URL (`permalink`) and the
  menu position (`nav_order`). A menu entry pointing at the posts
  section repositions the theme's blog page instead.
* **The home page.** `content/_index.md` becomes the body of al-folio's
  `about.md` landing page; a `{{< figure class="avatar" >}}` shortcode
  turns into the theme's profile picture.
* **Site identity.** Your name (from the home page title or
  `params.author`), `url` (from `baseURL`) and description are written
  into `_config.yml`, including jekyll-scholar's author-highlighting
  names.
* **Theme demo content is scrubbed**: al-folio's sample posts, news,
  projects, books, teachings and the Einstein pages are removed, so the
  site starts empty instead of generic.
* **`static/`** is copied to the site root, so absolute URLs like
  `/images/foo.png` keep working. With `only_referenced = TRUE` only
  files actually referenced by the migrated content (plus favicons) are
  copied -- old Hugo `static/` folders accumulate junk.
* **Shortcodes**: `{{< figure >}}` and `{{< youtube >}}` are converted;
  everything else is left in place and listed in the migration report.
  A figure's `class`/`width`/`height` become a kramdown attribute list,
  and the CSS rules behind those classes are ported from the Hugo
  theme's sass (variables resolved) into the new site's stylesheet, so
  images keep their original display size instead of rendering at
  natural size.
* **Pandoc image attributes** (`{width=100%}`) are rewritten as kramdown
  inline attribute lists (`{: width="100%"}`).

The report printed at the end lists what is left: posts with remaining
shortcodes and content *not* referenced by the Hugo menu (dead pages,
custom `layouts/`) -- review that list before assuming it can be dropped.

## Step 2 -- review the posts

Knit everything (no Jekyll needed) and read the output:

```{r}
build_site("~/path/to/my-new-site", local_jekyll = FALSE)
```

Things worth checking in `_posts/*.md`:

* Posts using **HTML widgets** (plotly, leaflet, ...) render as static
  Markdown here; widgets that relied on blogdown's HTML pipeline need
  rework (e.g. save a static image, or embed the widget's HTML manually).
* Shortcodes flagged in the report need a manual equivalent (a Liquid
  include, plain HTML, or just Markdown).
* Draft posts were kept with `published: false`; delete or finish them.
* Posts needing pandoc-only features -- citations (`bibliography:`),
  footnotes, cross-references -- can opt in with `knit_method: pandoc`
  in the front matter (needs rmarkdown + pandoc). New posts can also be
  written in Quarto (`new_post(format = "qmd")`), which brings those
  features natively via the Quarto CLI.

## Step 3 -- adapt the al-folio configuration

`migrate_hugo()` fills in your identity (name, `url`, description) but
al-folio's rich `_config.yml` deserves a pass by hand:

* Check `baseurl` (usually `""` for a user page) and browse the config
  for features you want (Google Analytics, comments, ...).
* Social icons live in `_data/socials.yml`. `migrate_hugo()` fills it
  from the profile links found in your Hugo pages (GitHub, LinkedIn,
  ORCID, Google Scholar, ResearchGate, StackOverflow, ...), plus your
  e-mail address and a CV/resume PDF link -- check the migration report
  for what was recognized and add anything missing by hand.
* Publications: with `publications = "bib"` (the default) the DOIs on
  the page are extracted and fetched as BibTeX from doi.org, so
  al-folio's jekyll-scholar renders the publications page with the
  theme's badges and buttons -- deterministic and reproducible (the DOI
  registry, not a heuristic parser or an LLM, is the source of truth).
  Publications without a DOI are listed in the report; add them to
  `_bibliography/papers.bib` by hand. `publications = "html"` migrates
  your hand-written page as-is instead -- that is also the automatic
  fallback when the page has no DOI or there is no network, but note
  that a page built around raw HTML (floats, flex `<div>`s) rarely
  survives kramdown intact.
* Accent color: al-folio defaults to purple. `theme_color = "red"` (or
  any palette name / hex value) sets it at migration time, and
  `set_theme_color()` changes it later -- it needs the theme gems, so on
  a first machine run `bundle_install()` and then `set_theme_color()`.

### Known al-folio gotchas

Verified with al-folio as of August 2026:

* **Jupyter notebooks.** al-folio refuses to build when the
  `jekyll-jupyter-notebook` plugin is enabled but `jupyter-nbconvert` is
  not on the `PATH`. Since a freshly migrated site has no `.ipynb`
  posts, `migrate_hugo()` removes the plugin from the `Gemfile` and
  from `_config.yml` for you. If you later want to write posts as
  notebooks, add both entries back and install Jupyter.
* **JavaScript minification.** Terser wants a JS runtime (Node.js). The
  build succeeds without one -- assets are just not minified -- so this
  only matters for production polish. GitHub Actions provides Node, so
  remote builds are unaffected.

## Step 4 -- preview locally (optional)

al-folio manages its plugins with Bundler, so install its gems once into
jekylldown's isolated environment, then serve:

```{r}
bundle_install("~/path/to/my-new-site")
serve_site("~/path/to/my-new-site")
# ... edit _source/*.Rmd; saving re-knits and reloads the browser
stop_server()
```

After `bundle_install()`, `build_site()` and `serve_site()` automatically
use `bundle exec jekyll`, so the theme's plugins are found.

## Step 5 -- deploy on GitHub Pages

1. Create a repository named `youruser.github.io` and push the new site
   to it (including `_source/` -- it is excluded from the Jekyll build but
   belongs in version control).
2. In the repository settings, enable GitHub Pages with **GitHub
   Actions** as the source: al-folio's bundled workflow
   (`.github/workflows/deploy.yml`) builds and publishes the site on
   every push.
3. Your R workflow from then on: `new_post()`, write, `build_site()` (or
   just knit via `build_site(local_jekyll = FALSE)`), commit, push. No
   more two-repository setup as with blogdown -- the source and the built
   site live in one repo, built remotely.

## Migration checklist

- [ ] `migrate_hugo()` ran clean; skipped posts (if any) handled
- [ ] posts knit with `build_site(local_jekyll = FALSE)`
- [ ] shortcodes from the report ported
- [ ] `_config.yml` reviewed (identity was set; check `baseurl`, extras)
- [ ] pages not in the Hugo menu (report's manual list) triaged
- [ ] `_data/socials.yml` reviewed (filled from your Hugo pages)
- [ ] publications page: migrated version kept, or
      `_bibliography/papers.bib` filled instead
- [ ] repo pushed, GitHub Actions build green
- [ ] old site archived (not deleted) once the new one is live
